Faster generic IND-CCA2 secure KEM using ``encrypt-then-MAC''

### 2024-10-08 etm-transform/manuscript/main.tex
Some words in conclusion

%%%% 1. DOCUMENTCLASS %%%%
%
% TODO: need more thorough references
% TODO: need better conclusion
% \documentclass[journal=tches,submission,notanonymous]{iacrtrans}
\documentclass[journal=tches,submission]{iacrtrans}
%%%% NOTES:
% - Change "journal=tosc" to "journal=tches" if needed
% - Change "submission" to "final" for final version
% - Add "notanonymous" to reveal authors
% - Add "spthm" for LNCS-like theorems

@@

\section{Introduction}\label{sec:introduction}
A key encapsulation mechanism (KEM) \cite{DBLP:journals/iacr/Shoup01} is a cryptographic primitive that allows two parties to establish a shared secret over an insecure channel. The desired security standard for a KEM is called indistinguishability under adaptive chosen-ciphertext attack (IND-CCA2). Intuitively speaking, IND-CCA2 security requires that no efficient adversary can distinguish a pseudorandom shared secret from truly random noise, even with access to a decapsulation oracle throughout the attack. However, building a provably IND-CCA2 KEM is tremendously difficult. Early attempts without formal proof, such as RSA encryption defined in PKCS\#1 v1.5 \cite{DBLP:journals/rfc/rfc2313}, were later shown to be vulnerable to practical chosen-ciphertext attack \cite{DBLP:conf/crypto/Bleichenbacher98}. In recent decades, the most viable approach has been to start with cryptographic primitives possessing weaker security properties, such as a public-key encryption (PKE) scheme with one-way security under chosen-plaintext attack (OW-CPA), then add steps to ensure ciphertext non-malleability \cite{bellare1998relations}. Some of the earliest proposals for generic IND-CCA2 secure constructions include OAEP \cite{bellare1995optimal}, Fujisaki-Okamoto transformation \cite{DBLP:conf/crypto/FujisakiO99}\cite{DBLP:journals/joc/FujisakiO13}, REACT \cite{DBLP:conf/ctrsa/OkamotoP01}, and GEM \cite{DBLP:conf/ctrsa/CoronHJPPT02}.

@@

\textbf{Correctness.} Following the definition in \cite{DBLP:conf/eurocrypt/DworkNR04} and \cite{hofheinz2017modular}, a $\pke$ is $\delta$-correct if:

@@

\section{Implementation}\label{sec:implementation}
ML-KEM is an \texttt{IND-CCA2} secure key encapsulation mechanism standardized by NIST in FIPS 203. The \texttt{IND-CCA2} security of ML-KEM is achieved in two steps. First, ML-KEM constructs an \texttt{IND-CPA} secure public key encryption scheme $\texttt{K-PKE}(\keygen, \encrypt, \decrypt)$ whose security is based on the conjectured intractability of the module learning with error (MLWE) problems against both classical and quantum adversaries. Then, the $U^{\not\bot}_m$ variant of the Fujisaki-Okamoto transformation \cite{hofheinz2017modular} is used to construct the KEM $\texttt{MLKEM}(\keygen, \encap, \decap)$ by calling $\texttt{K-PKE}(\keygen, \encrypt, \decrypt)$ as sub-routines. Because $\texttt{K-PKE}.\encrypt$ includes substantially more arithmetics than $\texttt{K-PKE}.\decrypt$, by using \textit{re-encryption} and \textit{de-randomization}, ML-KEM's decapsulation routine incurs significant computational cost.

We implemented the ``encrypt-then-MAC'' KEM construction using $\texttt{K-PKE}$ as the input PKE and compared its performance against ML-KEM under a variety of scenarios. The experimental data showed that while the ``encrypt-then-MAC'' construction adds a small amount of computational overhead to the encapsulation routine and a small increase in ciphertext size when compared with ML-KEM, it boasts enormous runtime savings in the decapsulation routine, which makes it particularly suitable for deployment in constrained environment. See appendix \ref{sec:ml-kem-and-kyber} for comparison with Kyber's third round submission to NIST's PQC competition.

@@

% TODO: How is ML-KEM different from Kyber round 3?

@@

    \begin{tabular}[t]{|cc|}
        \hline
        % TODO: this title needs to show that parameters are identical across all variants
        \multicolumn{2}{|c|}{\bf 128-bit security} \\
        \hline
        \multicolumn{2}{|c|}{size parameters (bytes)} \\
        $\pk$ size & 800 \\
        $\sk$ size & 1632 \\
        \text{ct} size & 768 \\
        \hline
        \multicolumn{2}{|c|}{$\keygen$ cycles/tick} \\
        Median & 75945 \\
        Average & 76171 \\
        \hline
    \end{tabular}
    \begin{tabular}[t]{|c|c|c|c|c|}
        \hline
        \multirow{2}{*}{KEM variant} 
        & \multicolumn{2}{|c|}{$\encap$ cycles/tick} 
        & \multicolumn{2}{|c|}{$\decap$ cycles/tick} \\
        \cline{2-5}
        & Median & Average & Median & Average \\
        \hline
        \text{ML-KEM-512} & 91467 & 92065 & 121185 & 121650 \\
        \hline
        \text{Kyber512} & 97811 & 98090 & 119937 & 120299 \\
        \hline
        % ML-KEM+-512
        $\text{ML-KEM}^+$-512 w/ Poly1305 & 93157 & 93626 & 33733 & 33908 \\
        \hline
        $\text{ML-KEM}^+$-512 w/ GMAC & 97369 & 97766 & 37725 & 37831 \\
        \hline
        $\text{ML-KEM}^+$-512 w/ CMAC & 99739 & 99959 & 40117 & 39943 \\
        \hline
        $\text{ML-KEM}^+$-512 w/ KMAC & 101009 & 101313 & 40741 & 40916 \\
        \hline
    \end{tabular}\vspace{0.3cm}

@@

\section{Conclusions and future works}\label{sec:future-works}
The ``encrypt-then-MAC'' transformation is a generic KEM construction that achieves IND-CCA2 security under the random oracle model if the input PKE is OW-PCA secure. Compared to the Fujisaki-Okamoto transformation, our construction replaced \emph{de-randomization} and \emph{re-encryption} with a message authenticator. At the cost of some minimal increase in communication size and encapsulation runtime, our construction achieves significant efficiency gains in the decapsulation routine. In practical key exchange protocols, our construction saves between 35-45\% in round trip time.

\textbf{Other suitable post-quantum instantiation.} In section \ref{sec:implementation}, we instantiated the ``encrypt-then-MAC'' KEM transformation with the PKE subroutines of ML-KEM. Unfortunately, ML-KEM is vulnerable to plaintext checking attacks that can recover complete secret key if the keypair is reused \cite{DBLP:conf/eurocrypt/BaetuDHTV19}. In fact, as Chris Peikert pointed out, because of the search-decision equivalence of LWE problems, almost all LWE-based cryptosystems will be vulerable to plaintext-checking attacks \cite{DBLP:conf/pqcrypto/Peikert14}. This naturally raises the question of finding other post-quantum primitives to instantiate ``encrypt-then-MAC'' with, such as other lattice-based cryptosystems not based on LWE, code-based cryptosystems and isogeny-based cryptosystems.

@@

\section{Appendix}\label{sec:appendix}
% TODO: brief introduction to the various authenticators should be in the appendix

\subsection{Performance comparison between $\texttt{ML-KEM}^+$ and Kyber}\label{sec:ml-kem-and-kyber}
ML-KEM directly evolved from CRYSTALS-Kyber's third round submission to NIST's post quantum cryptography competition. While their IND-CPA subroutines (see figure \ref{fig:k-pke-routines}) are identical, ML-KEM deviated from Kyber by choosing a different variant of the Fujisaki-Okamoto transformation.

    \begin{minipage}[t]{0.33\textwidth}
        \begin{algorithm}[H]
            \caption*{\texttt{K-PKE.KeyGen()}}
            \begin{algorithmic}[1]
                \State $A \leftsample R_q^{k \times k}$
                \State $\mathbf{s} \leftsample \mathcal{X}_{\eta_1}^k$
                \State $\mathbf{e} \leftsample \mathcal{X}_{\eta_1}^k$
                \State $\mathbf{t} \leftarrow A\mathbf{s} + \mathbf{e}$
                \State $\pk \leftarrow (A, \mathbf{t})$
                \State $\sk \leftarrow \mathbf{s}$
                \State \Return $(\pk, \sk)$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}\hfill
    \begin{minipage}[t]{0.33\textwidth}
        \begin{algorithm}[H]
            \caption*{\texttt{K-PKE.Enc(pk, m)}}
            \begin{algorithmic}[1]
                \Ensure $\pk = (A, \mathbf{t})$
                \Ensure $m \in R_2$
                \State $\mathbf{r} \leftsample \mathcal{X}_{\eta_1}^k$
                \State $\mathbf{e}_1 \leftsample \mathcal{X}_{\eta_2}^k$
                \State $e_2 \leftsample \mathcal{X}_{\eta_2}$
                \State $\mathbf{c}_1 \leftarrow A\mathbf{r} + \mathbf{e}_1$
                \State $c_2 \leftarrow \mathbf{t}^\intercal \mathbf{r} + e_2 + m \cdot \lfloor \frac{q}{2}\rceil$
                \State \Return $(\mathbf{c}_1, c_2)$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}\hfill
    \begin{minipage}[t]{0.3\textwidth}
        \begin{algorithm}[H]
            \caption*{\texttt{K-PKE.Dec(sk, c)}}
            \begin{algorithmic}[1]
                \Ensure $c = (\mathbf{c}_1, c_2)$
                \Ensure $\sk = \mathbf{s}$
                \State $\hat{m} \leftarrow c_2 - \mathbf{c}_1^\intercal \cdot \mathbf{s}$
                \State $\hat{m} \leftarrow \texttt{Round}(\hat{m})$
                \State \Return $\hat{m}$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \caption{\texttt{K-PKE} routines are identical between Kyber and ML-KEM}\label{fig:k-pke-routines}
\end{figure}

CRYSTALS-Kyber uses the $U^{\not\bot}$ variant, where the shared secret is derived from both the plaintext and the ciphertext. On the other hand, because by using \emph{re-encryption} and \emph{de-randomization}, the PKE is already made \emph{rigid}, the CRYSTALS-Kyber team decided to use the $U^{\not\bot}_m$ variant, where the shared secret is derived from the plaintext alone.

\begin{figure}[H]
    \centering
    \begin{minipage}[t]{0.45\textwidth}
        \begin{algorithm}[H]
            \caption*{\texttt{KEM.KeyGen()}}
            \begin{algorithmic}[1]
                \State $z \leftsample \{0,1\}^{256}$
                \State $(\pk, \sk^\prime) \leftsample \texttt{PKE.KeyGen}()$
                \State $\sk \leftarrow (\sk^\prime \Vert \pk \Vert H(\pk) \Vert z)$
                \State \Return $(\pk, \sk)$
            \end{algorithmic}
        \end{algorithm}
        \begin{algorithm}[H]
            \caption*{\texttt{KEM.Encap(pk)}}
            \begin{algorithmic}[1]
                \State $m \leftsample \{0,1\}^{256}$
                \State $(\overline{K}, r) \leftarrow G(m \Vert H(\pk))$
                \State $c \leftarrow \texttt{PKE.Enc}(\pk, m, r)$
                \State $K \leftarrow \texttt{KDF}(\overline{K} \Vert H(c))$
                    \Comment{$U^{\not\bot}$}
                \State $K \leftarrow \overline{K}$
                    \Comment{$U^{\not\bot}_m$}
                \State \Return $(c, K)$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \begin{minipage}[t]{0.5\textwidth}
        \begin{algorithm}[H]
            \caption*{\texttt{KEM.Decap(sk, c)}}
            \begin{algorithmic}[1]
                \Ensure $\sk = (\sk^\prime \Vert \pk \Vert H(\pk) \Vert z)$
                \State $\hat{m} \leftarrow \texttt{PKE.Dec}(\sk^\prime, c)$
                \State $(\overline{K}, \hat{r}) \leftarrow G(\hat{m} \Vert H(\pk))$
                \If{$\texttt{PKE.Enc}(\pk, \hat{m}, \hat{r}) = c$}
                    \State $K \leftarrow \texttt{KDF}(\overline{K}, H(c))$
                        \Comment{$U^{\not\bot}$}
                    \State $K \leftarrow \overline{K}$
                        \Comment{$U^{\not\bot}_m$}
                \Else 
                    \State $K \leftarrow \texttt{KDF}(z \Vert H(c))$
                \EndIf
                \State \Return $K$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}

    \caption{
        Kyber uses $U^{\not\bot}$ variant. ML-KEM uses $U^{\not\bot}_m$ variant.
    }\label{fig:kyber-ml-kem-compare}
\end{figure}

The reason for ML-KEM to use a different variant of the Fujisaki-Okamoto transformation is two-fold. The first reason is performance: using the $U^{\not\bot}_m$ transformation saves the need to hash the ciphertext, and since Kyber/ML-KEM's performance is mainly bottlenecked by the symmetric components, omitting the hash leads to significant runtime savings (up to 17\% in AVX-2 optimized implementations). The second reason is the simplified security proof and tighter security bounds of the $U^{\not\bot}_m$ variant compared to the $U^{\not\bot}$ variant. We will omit the details of the security proof and refer readers to \cite{hofheinz2017modular}.

In section \ref{sec:implementation}, we mainly compared $\texttt{ML-KEM}^+$ with \texttt{ML-KEM}, but the we would like to point out that, because Kyber uses the $U^{\not\bot}$ variant and needs to hash the ciphertext for deriving the shared secret, the performance advantage of $\texttt{ML-KEM}^+$ over Kyber will be even greater.

>>>>

%%%% 1. DOCUMENTCLASS %%%%
%
% \documentclass[journal=tches,submission,notanonymous]{iacrtrans}
\documentclass[journal=tches,submission]{iacrtrans}
%%%% NOTES:
% - Change "journal=tosc" to "journal=tches" if needed
% - Change "submission" to "final" for final version
% - Add "notanonymous" to reveal authors
% - Add "spthm" for LNCS-like theorems

@@

\section{Introduction}\label{sec:introduction}
A key encapsulation mechanism (KEM) \cite{DBLP:journals/iacr/Shoup01} is a cryptographic primitive that allows two parties to establish a shared secret over an insecure channel. The desired security standard for a KEM is called indistinguishability under adaptive chosen-ciphertext attack (IND-CCA2). Intuitively speaking, IND-CCA2 security requires that no efficient adversary can distinguish a pseudorandom shared secret from truly random noise, even with access to a decapsulation oracle throughout the attack. However, building a provably IND-CCA2 KEM is tremendously difficult. Early attempts without formal proof, such as RSA encryption defined in PKCS\#1 v1.5 \cite{DBLP:journals/rfc/rfc2313}, were later shown to be vulnerable to practical chosen-ciphertext attack \cite{DBLP:conf/crypto/Bleichenbacher98}. In recent decades, the most viable approach has been to start with cryptographic primitives possessing weaker security properties, such as a public-key encryption (PKE) scheme with one-way security under chosen-plaintext attack (OW-CPA), then add steps to ensure ciphertext non-malleability \cite{DBLP:conf/asiacrypt/BellareN00}. Some of the earliest proposals for generic IND-CCA2 secure constructions include OAEP \cite{DBLP:conf/eurocrypt/BellareR94}, Fujisaki-Okamoto transformation \cite{DBLP:conf/crypto/FujisakiO99}\cite{DBLP:journals/joc/FujisakiO13}, REACT \cite{DBLP:conf/ctrsa/OkamotoP01}, and GEM \cite{DBLP:conf/ctrsa/CoronHJPPT02}.

@@

\textbf{Correctness.} Following the definition in \cite{DBLP:conf/eurocrypt/DworkNR04}, a $\pke$ is $\delta$-correct if:

@@

\section{Implementation}\label{sec:implementation}
ML-KEM is an \texttt{IND-CCA2} secure key encapsulation mechanism standardized by NIST in FIPS 203. The \texttt{IND-CCA2} security of ML-KEM is achieved in two steps. First, ML-KEM constructs an \texttt{IND-CPA} secure public key encryption scheme $\texttt{K-PKE}(\keygen, \encrypt, \decrypt)$ whose security is based on the conjectured intractability of the module learning with error (MLWE) problems against both classical and quantum adversaries. Then, the $U^{\not\bot}_m$ variant of the Fujisaki-Okamoto transformation is used to construct the KEM $\texttt{MLKEM}(\keygen, \encap, \decap)$ by calling $\texttt{K-PKE}(\keygen, \encrypt, \decrypt)$ as sub-routines. Because $\texttt{K-PKE}.\encrypt$ includes substantially more arithmetics than $\texttt{K-PKE}.\decrypt$, by using \textit{re-encryption} and \textit{de-randomization}, ML-KEM's decapsulation routine incurs significant computational cost.

We implemented the ``encrypt-then-MAC'' KEM construction using $\texttt{K-PKE}$ as the input PKE and compared its performance against ML-KEM under a variety of scenarios. The experimental data showed that while the ``encrypt-then-MAC'' construction adds a small amount of computational overhead to the encapsulation routine and a small increase in ciphertext size when compared with ML-KEM, it boasts enormous runtime savings in the decapsulation routine, which makes it particularly suitable for deployment in constrained environment.

@@

@@

    \begin{tabular}[t]{|cc|}
        \hline
        \multicolumn{2}{|c|}{\bf 128-bit security} \\
        \hline
        \multicolumn{2}{|c|}{size parameters (bytes)} \\
        $\pk$ size & 800 \\
        $\sk$ size & 1632 \\
        \text{ct} size & 768 \\
        \hline
        \multicolumn{2}{|c|}{$\keygen$ cycles/tick} \\
        Median & 75945 \\
        Average & 76171 \\
        \hline
    \end{tabular}
    \begin{tabular}[t]{|c|c|c|c|c|}
        \hline
        \multirow{2}{*}{KEM variant} 
        & \multicolumn{2}{|c|}{$\encap$ cycles/tick} 
        & \multicolumn{2}{|c|}{$\decap$ cycles/tick} \\
        \cline{2-5}
        & Median & Average & Median & Average \\
        \hline
        \text{ML-KEM-512} & 91467 & 92065 & 121185 & 121650 \\
        \hline
        \text{Kyber512} & 97811 & 98090 & 119937 & 120299 \\
        \hline
        % ML-KEM+-512
        $\text{ML-KEM}^+$-512 w/ Poly1305 & 93157 & 93626 & 33733 & 33908 \\
        \hline
        $\text{ML-KEM}^+$-512 w/ GMAC & 97369 & 97766 & 37725 & 37831 \\
        \hline
        $\text{ML-KEM}^+$-512 w/ CMAC & 99739 & 99959 & 40117 & 39943 \\
        \hline
        $\text{ML-KEM}^+$-512 w/ KMAC & 101009 & 101313 & 40741 & 40916 \\
        \hline
    \end{tabular}\vspace{0.3cm}

@@

\section{Conclusions and future works}\label{sec:future-works}
The ``encrypt-then-MAC'' transformation is a generic KEM construction whose IND-CCA2 security reduces to the OW-PCA security of the input PKE and the one-time existential unforgeability of the input MAC in the random oracle model. Compared to the Fujisaki-Okamoto transformation, the ``encrypt-then-MAC'' replaces the computationally expensive re-encryption with computing a MAC tag. At the cost of minimal increase in encapsulation cost and ciphertext size, the ``encrypt-then-MAC'' substantially improves the efficiency of the decapsulation routine. Where the input PKE's encryption is slower than decryption, the ``encrypt-then-MAC'' KEM achieves meaningful time savings in practical key exchange protocols.

Unfortunately, ML-KEM is not OW-PCA secure. In fact, as Chris Peikert pointed out in \cite{DBLP:conf/pqcrypto/Peikert14}, most lattice-based cryptosystems are not OW-PCA secure due to the search-decision equivalence of lattice problems. While we instantiated ``encrypt-then-MAC'' with ML-KEM subroutines for performance comparison, the resulting KEM is not chosen-ciphertext secure, and should not be used in production systems. The natural questions is thus to find a suitable cryptosystem to instantiate ``encrypt-then-MAC'' with.

\textbf{RSA and ElGamal.} Combining ``encrypt-then-MAC'' with ElGamal results in the ``Hashed ElGamal'' scheme proposed in \cite{DBLP:journals/iacr/AbdallaBR99}\cite{DBLP:conf/ctrsa/AbdallaBR01}. RSA is also known to be OW-PCA secure because it is a trapdoor permutation and thus a rigid PKE. However, because of RSA's rigidity, there exists even more efficient KEM transformation (such as RSA-KEM \cite{DBLP:journals/iacr/Shoup01}) that only adds a single hash to the base PKE routines. While applying ``encrypt-then-MAC'' to RSA will result in an IND-CCA2 secure KEM, such construction offers no meaningful advantage to RSA-KEM.

\textbf{Code-based cryptography.} Because the general problem of decoding a linear code is proven NP hard \cite{DBLP:journals/tit/BerlekampMT78}, code-based cryptosystems may not suffer from the inherent search-decision equivalence of lattice problems and thus be viable candidates with OW-PCA security. Unfortunately, among the code-based submissions to NIST PQC, HQC \cite{melchor2018hamming} and BIKE \cite{aragon2022bike} are known to be vulnerable to key-recovery plaintext-checking attacks (KR-PCA) \cite{DBLP:journals/tches/TanakaUXITH23}. On the other hand, classic McEliece \cite{classicmceliecespec} seems to be PCA secure and thus a viable candidate, though in classic McEliece, the decoding routine is more expensive than the encryption routine, so applying ``encrypt-then-MAC'' may not yield meaningful performance gains.

\textbf{Isogeny-based cryptography.} The intractability assumptions of isogeny-based cryptography resemble the classical Diffie-Hellman assumptions, and it seems possible to formulate a ``Gap Diffie-Hellman assumption'' in supersingular isogeny \cite{DBLP:conf/icisc/FujiokaTTY18}. While SIKE and SIDH were found to be insecure \cite{DBLP:conf/eurocrypt/CastryckD23}, other isogeny-based cryptosystem such as CSIDH \cite{DBLP:conf/asiacrypt/Castryck0MPR18} remains unaffected by the aforementioned attack and might be suitable candidates.

@@

% \section{Appendix}\label{sec:appendix}
% % TODO: brief introduction to the various authenticators should be in the appendix

% \subsection{Performance comparison between $\texttt{ML-KEM}^+$ and Kyber}\label{sec:ml-kem-and-kyber}
% ML-KEM directly evolved from CRYSTALS-Kyber's third round submission to NIST's post quantum cryptography competition. While their IND-CPA subroutines (see figure \ref{fig:k-pke-routines}) are identical, ML-KEM deviated from Kyber by choosing a different variant of the Fujisaki-Okamoto transformation.

% \begin{figure}[H]
%     \centering

%     \begin{minipage}[t]{0.33\textwidth}
%         \begin{algorithm}[H]
%             \caption*{\texttt{K-PKE.KeyGen()}}
%             \begin{algorithmic}[1]
%                 \State $A \leftsample R_q^{k \times k}$
%                 \State $\mathbf{s} \leftsample \mathcal{X}_{\eta_1}^k$
%                 \State $\mathbf{e} \leftsample \mathcal{X}_{\eta_1}^k$
%                 \State $\mathbf{t} \leftarrow A\mathbf{s} + \mathbf{e}$
%                 \State $\pk \leftarrow (A, \mathbf{t})$
%                 \State $\sk \leftarrow \mathbf{s}$
%                 \State \Return $(\pk, \sk)$
%             \end{algorithmic}
%         \end{algorithm}
%     \end{minipage}\hfill
%     \begin{minipage}[t]{0.33\textwidth}
%         \begin{algorithm}[H]
%             \caption*{\texttt{K-PKE.Enc(pk, m)}}
%             \begin{algorithmic}[1]
%                 \Ensure $\pk = (A, \mathbf{t})$
%                 \Ensure $m \in R_2$
%                 \State $\mathbf{r} \leftsample \mathcal{X}_{\eta_1}^k$
%                 \State $\mathbf{e}_1 \leftsample \mathcal{X}_{\eta_2}^k$
%                 \State $e_2 \leftsample \mathcal{X}_{\eta_2}$
%                 \State $\mathbf{c}_1 \leftarrow A\mathbf{r} + \mathbf{e}_1$
%                 \State $c_2 \leftarrow \mathbf{t}^\intercal \mathbf{r} + e_2 + m \cdot \lfloor \frac{q}{2}\rceil$
%                 \State \Return $(\mathbf{c}_1, c_2)$
%             \end{algorithmic}
%         \end{algorithm}
%     \end{minipage}\hfill
%     \begin{minipage}[t]{0.3\textwidth}
%         \begin{algorithm}[H]
%             \caption*{\texttt{K-PKE.Dec(sk, c)}}
%             \begin{algorithmic}[1]
%                 \Ensure $c = (\mathbf{c}_1, c_2)$
%                 \Ensure $\sk = \mathbf{s}$
%                 \State $\hat{m} \leftarrow c_2 - \mathbf{c}_1^\intercal \cdot \mathbf{s}$
%                 \State $\hat{m} \leftarrow \texttt{Round}(\hat{m})$
%                 \State \Return $\hat{m}$
%             \end{algorithmic}
%         \end{algorithm}
%     \end{minipage}
%     \caption{\texttt{K-PKE} routines are identical between Kyber and ML-KEM}\label{fig:k-pke-routines}
% \end{figure}

% CRYSTALS-Kyber uses the $U^{\not\bot}$ variant, where the shared secret is derived from both the plaintext and the ciphertext. On the other hand, because by using \emph{re-encryption} and \emph{de-randomization}, the PKE is already made \emph{rigid}, the CRYSTALS-Kyber team decided to use the $U^{\not\bot}_m$ variant, where the shared secret is derived from the plaintext alone.

% \begin{figure}[H]
%     \centering
%     \begin{minipage}[t]{0.45\textwidth}
%         \begin{algorithm}[H]
%             \caption*{\texttt{KEM.KeyGen()}}
%             \begin{algorithmic}[1]
%                 \State $z \leftsample \{0,1\}^{256}$
%                 \State $(\pk, \sk^\prime) \leftsample \texttt{PKE.KeyGen}()$
%                 \State $\sk \leftarrow (\sk^\prime \Vert \pk \Vert H(\pk) \Vert z)$
%                 \State \Return $(\pk, \sk)$
%             \end{algorithmic}
%         \end{algorithm}
%         \begin{algorithm}[H]
%             \caption*{\texttt{KEM.Encap(pk)}}
%             \begin{algorithmic}[1]
%                 \State $m \leftsample \{0,1\}^{256}$
%                 \State $(\overline{K}, r) \leftarrow G(m \Vert H(\pk))$
%                 \State $c \leftarrow \texttt{PKE.Enc}(\pk, m, r)$
%                 \State $K \leftarrow \texttt{KDF}(\overline{K} \Vert H(c))$
%                     \Comment{$U^{\not\bot}$}
%                 \State $K \leftarrow \overline{K}$
%                     \Comment{$U^{\not\bot}_m$}
%                 \State \Return $(c, K)$
%             \end{algorithmic}
%         \end{algorithm}
%     \end{minipage}
%     \begin{minipage}[t]{0.5\textwidth}
%         \begin{algorithm}[H]
%             \caption*{\texttt{KEM.Decap(sk, c)}}
%             \begin{algorithmic}[1]
%                 \Ensure $\sk = (\sk^\prime \Vert \pk \Vert H(\pk) \Vert z)$
%                 \State $\hat{m} \leftarrow \texttt{PKE.Dec}(\sk^\prime, c)$
%                 \State $(\overline{K}, \hat{r}) \leftarrow G(\hat{m} \Vert H(\pk))$
%                 \If{$\texttt{PKE.Enc}(\pk, \hat{m}, \hat{r}) = c$}
%                     \State $K \leftarrow \texttt{KDF}(\overline{K}, H(c))$
%                         \Comment{$U^{\not\bot}$}
%                     \State $K \leftarrow \overline{K}$
%                         \Comment{$U^{\not\bot}_m$}
%                 \Else 
%                     \State $K \leftarrow \texttt{KDF}(z \Vert H(c))$
%                 \EndIf
%                 \State \Return $K$
%             \end{algorithmic}
%         \end{algorithm}
%     \end{minipage}

%     \caption{
%         Kyber uses $U^{\not\bot}$ variant. ML-KEM uses $U^{\not\bot}_m$ variant.
%     }\label{fig:kyber-ml-kem-compare}
% \end{figure}

% The reason for ML-KEM to use a different variant of the Fujisaki-Okamoto transformation is two-fold. The first reason is performance: using the $U^{\not\bot}_m$ transformation saves the need to hash the ciphertext, and since Kyber/ML-KEM's performance is mainly bottlenecked by the symmetric components, omitting the hash leads to significant runtime savings (up to 17\% in AVX-2 optimized implementations). The second reason is the simplified security proof and tighter security bounds of the $U^{\not\bot}_m$ variant compared to the $U^{\not\bot}$ variant. We will omit the details of the security proof and refer readers to \cite{hofheinz2017modular}.

% In section \ref{sec:implementation}, we mainly compared $\texttt{ML-KEM}^+$ with \texttt{ML-KEM}, but the we would like to point out that, because Kyber uses the $U^{\not\bot}$ variant and needs to hash the ciphertext for deriving the shared secret, the performance advantage of $\texttt{ML-KEM}^+$ over Kyber will be even greater.
